Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.

### 2022-06-09 manuscript.Rmd
use apa citation style

---
title: "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization."
output:
  bookdown::pdf_document2:
    latex_engine: "xelatex"
    toc: false
    number_sections: false
header-includes:
   - \usepackage{libertine}
fontsize: 11pt
linestretch: 1.2
geometry: "left=3.9cm, right=3.3cm, top=2.5cm, bottom=3cm"
papersize: a4
bibliography: references.bib
abstract: "`r paste0(readLines('abstract.md'), collapse = ' ')`"
---

@@

A successful prediction should lend more credibility to a theory than a postdiction, all else equal.
The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [see @hoyningen-hueneContextDiscoveryContext2006].
Confusing exploratory findings with confirmed theories has lead to a crisis of confidence in the results of empirical sciences [@ioannidisWhyMostPublished2005] and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
As a response, more and more researchers preregister their studies [@nosekPreregistrationRevolution2018] to honour this distinction and produce results that are considered confirmatory.
Indeed, rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results produced by an arduous process of trial and error [@wagenmakersAgendaPurelyConfirmatory2012].
However, while many researchers intuitively recognize this as an advantage of preregistration, some remain unconvinced.
These skeptics question why a study should be preregistered at all.
Even researchers convinced of the advantages of preregistration face difficult problems when they exclusively rely on an intuitive appeal.
Specifially, an intuitive convinction is unable to address pragmatic questions, such as what to include, in how much detail in the preregistration, or about the conditions that warrant deviating from it.
If the research community remains confused about the purpose of preregistration, it is bound to misapply it and hence decrease the efficiency of the scientific endeavour. 
We thus propose a principled justification for preregistration with a formal objective against which researchers can evaluate several pragmatic trade-offs.

The scientific community asserts that the objective of preregistration is to distinguish confirmatory from exploratory research [@mellorEasyPreregistrationWill2018; @nosekPreregistrationRevolution2018; @wagenmakersAgendaPurelyConfirmatory2012].
Taken at face value, this objective implies that research is confirmatory if, and only if, it is preregistered.
However, this postulate is not warrented.
Researchers can conduct confirmatory research without preregisteration, though it might be difficult to convince other researchers of the confirmatory nature of their research.
The converse, therefore preregistered but not strictly confirmatory studies are also commonplace.
Researchers may apply two strategies to evade the self-imposed restrictions of preregistrations.
One strategy is to write a loose preregistration, to begin with; another is to deviate from the preregistration afterwards.
Both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating desirable results no matter the nature of the phenomenon under study.
Insisting on equating preregistration and confirmation has, hence, led to criticism that preregistration is actually harmfull all things considered, and neither sufficient nor necessary to establish confirmation [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021]

These issues arise from a fundamental confusion about the objective of preregistration.
Researchers are bound to be confused when they delegate an important scientific judgment to a simple decision rule, that only takes into account the existence of a preregistration.
Preregistering an inherently exploratory analysis (like testing dozens of relations) does not make it confirmatory, nor will a carefully conducted confirmatory study become exploratory if the researcher deviates from the preregistration in minor details.
Equating confirmatory research with preregistered research is only possible for studies that expect low type I error rate and will run/be analyzed without changes no matter what [@simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
Under these conditions, the rule holds, but such restricted use makes preregistration a niche solution unable to match the greater problem of replicability in psychology and elsewhere.

We show that the simple decision rule is just a special case of a more general conceptualization under Bayesian reasoning. 
To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [we borrow the term from @meehlTheoreticalRisksTabular1978], which is inversely related to type I error rate.
We then outline two interpretations of how theoretical risk is impacted by preregistration.
The first interpretation corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently by limiting type I error.
The second interpretation is our main contribution and demonstrates the broad applicability of preregistration for both exploratory and confirmatory studies that are implemented as preregistered or have undergone changes after preregistration.
Following this interpretation, the theoretical risk is not necessarily directly maximized by preregistration, but rather the uncertainty in judging the theoretical risk is minimized.

To arrive at this interpretation, we rely on three arguments.
The first is that theoretical risk is vital for judging theories.
The second argument is that the theoretical risk for a given study is generally uncertain.
The third and last argument is that this uncertainty is reduced by applying preregistration.
We conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for or against a theory, preregistration is warranted for any study.

@@

Let us start by defining what we call expected epistemic value.
If researchers plan to conduct a study, they usually hope it will change their assessment of some theory's verisimilitude (truthlikeness).
In other words, they hope to learn something from conducting the study.
The amount of knowledge researchers gain from a particular study concerning the verisimilitude of a specific theory is what we call epistemic value.
While researchers can not know what exactly they will learn from a study, they can form an expectation that helps them decide which study to conduct.
This expectation is what we term expected epistemic value.
To make our three arguments, we must assume three things about this estimation process and how it relates to choosing a study to conduct.

@@

The assumption of rationality can be connected to Bayesian reasoning and leads to our adoption of the framework.
Our rationale is as follows.
Researchers who decide to conduct a study are akin to choosing a study to bet on.
They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch Book" argument [@christensenCleverBookiesCoherent1991].
This argument states that any better must follow the axioms of probability to avoid being "irrational", i.e., accepting bets that lead to sure losses.
Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modelling what liberties researchers have in choosing a study.
Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
The axioms of probability are sufficient to derive the Bayes formula, on which we will heavily rely for our further arguments.
The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of probability; that remains a leap of faith.
Please note that our decision to adopt this aspect of the Bayesian philosophy of science does not imply anything about the statistical methods researchers use.
In fact, this conceptualization is purposefully reductionistic to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.

>>>>

---
title: "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization."
output:
  bookdown::pdf_document2:
    latex_engine: "xelatex"
    toc: false
    number_sections: false
header-includes:
   - \usepackage{libertine}
fontsize: 11pt
linestretch: 1.2
geometry: "left=3.9cm, right=3.3cm, top=2.5cm, bottom=3cm"
papersize: a4
bibliography: references.bib
abstract: "`r paste0(readLines('abstract.md'), collapse = ' ')`"
csl: apa.csl
---

@@

A successful prediction should lend more credibility to a theory than a postdiction, all else equal.
Much has been theorized on the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [see @hoyningen-hueneContextDiscoveryContext2006].
To honour this distinction and produce results that are considered confirmatory, many researchers opt to preregister their studies [@nosekPreregistrationRevolution2018].
Indeed, rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results produced by an arduous process of trial and error [@wagenmakersAgendaPurelyConfirmatory2012].
The scientific community widely agrees that actively concealing such "trying for results" is misconduct --- but how exactly is it impeding scientific progress?
We believe that a rigorous answer to this question enables us to clarify common misunderstandings that lead researchers to deem preregistration as unfit for their purposes.
These misunderstandings are interesting from a conceptual perspective but also lead to decreased efficiency of the scientific process if researchers misguidedly decide that preregistration is not applicable for their intended study.
Asking researchers to invest time and resources into preregistration demands a principled justification of its utility.

One reason why researchers are hesitant to preregister is that they feel restricted by the boundaries specified in their preregistration [@goldin-meadowWhyPreregistrationMakes2016].
Even researchers who have been persuaded of its advantages sometimes apply strategies to skirt these self-imposed restrictions.
One strategy is to write a loose preregistration, to begin with; another is to deviate from the preregistration afterwards.
To complicate matters further, both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating desirable results no matter the nature of the phenomenon under study.
To summarize, the conundrum is that the current conceptualization of preregistration deems a study "properly" preregistered even if it contains a recipe for p-hacking before data collection but "improper" if the researcher did technically deviate afterwards.
This, of course, goes against most researchers' intuition about what is considered exploratory and confirmatory.
We take these issues as signals of a misunderstanding about how preregistration contributes to scientific progress.
In particular, we want to challenge the common view that preregistration somehow cleanly separates confirmatory from exploratory.

Deciding if a study is confirmatory or exploratory is a judgment made by the researcher.
The issues outlined above arise by delegating this judgment to a simple decision rule.
That is, if something is preregistered, it is considered confirmatory, and if it is not preregistered, it is considered exploratory.
This rule of thumb is often applied implicitly.
Other times it is stated as the explicit goal of preregistration [@mellorEasyPreregistrationWill2018].
In both cases, it leads to confusion when applied outside specific bounds.
Preregistering an inherently exploratory analysis (like testing dozens of relations) does not make it confirmatory, nor will a carefully conducted confirmatory study become exploratory if the researcher deviates from the preregistration in minor details.
To uphold the simplicity of the decision rule, it is routinely suggested only to preregister when a study is expected to have a low type I error rate by design and will run/be analyzed without changes no matter what [@simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
Under these conditions, the rule holds, but such restricted use makes preregistration a niche solution unable to match the greater problem of replicability in psychology and elsewhere.

We show that this simple decision rule is a special case of a more general conceptualization under Bayesian reasoning.
To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [we borrow the term from @meehlTheoreticalRisksTabular1978], which is inversely related to type I error rate.
We then outline two interpretations of how theoretical risk is impacted by preregistration.
The first interpretation corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently by limiting type I error.
The second interpretation is our main contribution and demonstrates the broad applicability of preregistration for both exploratory and confirmatory studies that are implemented as preregistered or have undergone changes after preregistration.
Following this interpretation, the theoretical risk is not necessarily directly maximized by preregistration, but rather the uncertainty in judging the theoretical risk is minimized.

To arrive at this interpretation, we rely on three arguments.
The first is that theoretical risk is vital for judging theories.
This argument comes natural to a (Neo‑)Popperian or falsificationist, but  is not self-evident for a conformationist.
The second argument is that the theoretical risk for a given study is generally uncertain.
The third and last assumption is that this uncertainty is reduced by applying preregistration.

When we put these arguments into the Bayes' formula, we conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for or against a theory, preregistration is warranted for any study.
We will later discuss how our argument relates to a (Neo‑)Popperian philosophy of science.
For now, we stick to a Bayesian rationale since quantifying uncertainty is crucial for our argument, which would require considerable thought in a falsificationist framework.

@@

Let us start by defining what we call expected epistemic value.
If researchers plan to conduct a study, they usually hope it will change their assessment of some theory's verisimilitude (truthlikeness).
In other words, they hope to learn something from conducting the study.
The amount of knowledge researchers gain from a particular study concerning the verisimilitude of a specific theory is what we call epistemic value.
While researchers can not know what exactly they will learn from a study, they can form an expectation that helps them decide which study to conduct.
This expectation is what we term expected epistemic value.
We assume three things about this estimation process and how it relates to choosing a study to conduct.

@@

The first assumption leads to our adoption of a Bayesian framework.
Our rationale is as follows.
Researchers who decide to conduct a study are akin to choosing a study to bet on.
They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch Book" argument [@christensenCleverBookiesCoherent1991].
This argument states that any better must follow the axioms of probability to avoid being "irrational", i.e., accepting bets that lead to sure losses.
Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modelling what liberties researchers have in choosing a study.
Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
The axioms of probability are sufficient to derive the Bayes formula, on which we will heavily rely for our further arguments.
The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of probability; that remains a leap of faith.
Please note that our decision to adopt this aspect of the Bayesian philosophy of science does not imply anything about the statistical methods researchers use.
In fact, this conceptualization is purposefully reductionistic to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
